### 2022-07-31 latex/Chapters/Examples.tex
update

Unlike the Deep Learner, the Deep Reinforcement Learner learns unsupervised (in the sense that there is no need to pre-solve puzzles to tell if what the desired costs are), since it generates its own target using a combination of a target network and the simple min rule described in \ref{sec:codelearners}. Let us see here how to run it on the (n=4, m=2) SP for instance. The following code snipet will run a Deep Reinforcement Learner for a maximum of 25,000 epochs, generating randomly 10 sequences of puzzles shuffled 50 times from goal state every time it updates the target network. That update will happen either after 500 epochs or when the (MSE) loss on the value function gets under one thousands of the max target value. The learner will stop if it reaches 25,000 epochs or if the target network updates 40 times. The network trained is a fully connected network with 3 hidden layers and the puzzles are one-hot encoded. We use the same optimiser and scheduler as in the previous subsection.

@@

\textbf{manhattan heuristic}

\textbf{perfect heuristic}

\textbf{deep learning heuristic}

\textbf{deep reinforcement learning heuristic}

>>>>

Unlike the Deep Learner, the Deep Reinforcement Learner learns unsupervised (in the sense that there is no need to pre-solve puzzles to tell if what the actual (target) costs are), since it generates its own target using a combination of a target network and the simple min rule described in \ref{sec:codelearners}. Let us see here how to run it on the (n=4, m=2) SP for instance. The following code snipet will run a Deep Reinforcement Learner for a maximum of 25,000 epochs, generating randomly 10 sequences of puzzles shuffled 50 times from goal state every time it updates the target network. That update will happen either after 500 epochs or when the (MSE) loss on the value function gets under one thousands of the max target value. The learner will stop if it reaches 25,000 epochs or if the target network updates 40 times. The network trained is a fully connected network with 3 hidden layers and the puzzles are one-hot encoded. We use the same optimiser and scheduler as in the previous subsection.

@@

\subsubsection{Manhattan heuristic}

\afblue
\paragraph{}{\textbf{python code -- depth first search solver}}
\begin{python}
####################################################################
from rubiks.heuristics.heuristic import Heuristic
from rubiks.puzzle.puzzle import Puzzle
from rubiks.solvers.solver import Solver
####################################################################
if '__main__' == __name__:
    puzzle_type = Puzzle.sliding_puzzle
    tiles=[[3, 8, 6], [4, 1, 5], [0, 7, 2]]
    solver_type=Solver.astar
    heuristic_type=Heuristic.manhattan
    plus=False
    action_type=Solver.do_solve
    print(Solver.factory(**globals()).action().to_str_light())
####################################################################
\end{python}
\black
We run this example twice, one with simple Manhattan and one with Manhattan++. As can be seen from the output below, the Manhattan++ improves on the number of expanded nodes (as expected, since the heuristic is less optimistic while retaining its optimality property). As we have seen before, the overhead of adding penalty for the linear constraint (which have all been precomputed and stored in a database) is more than compensated for by the reduction in nodes expansion!

\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{./Figures/exampleastarmanhattansolver}
%\decoRule
\caption[Examples]{a* manhattan solver example}
\label{fig:exampleastarmanhattansolver}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{./Figures/exampleastarmanhattanplussolver}
%\decoRule
\caption[Examples]{a* manhattan++ solver example}
\label{fig:exampleastarmanhattanplussolver}
\end{figure}

\subsubsection{Perfect Heuristic}
To run A{*} with a perfect heuristic, we just need to specify the heuristic type as such, and set the parameter \textit{model\_file\_name} to point to a pre-recorded database populated by the PerfectLearner (see earlier section \ref{PLSS}).

\afblue
\paragraph{}{\textbf{python code -- depth first search solver}}
\begin{python}
####################################################################
from rubiks.heuristics.heuristic import Heuristic
from rubiks.puzzle.puzzle import Puzzle
from rubiks.solvers.solver import Solver
from rubiks.utils.utils import get_model_file_name
####################################################################
if '__main__' == __name__:
    puzzle_type = Puzzle.sliding_puzzle
    n=3
    nb_shuffles=40
    solver_type=Solver.astar
    heuristic_type=Heuristic.perfect
    model_file_name = get_model_file_name(puzzle_type=puzzle_type,
                                          dimension=(n, n),
                                          model_name=Heuristic.perfect)
    action_type=Solver.do_solve
    print(Solver.factory(**globals()).action().to_str_light())
####################################################################
\end{python}
\black
Running this code snipet will output something like (the shuffle is random obviously):

\begin{figure}[H]
\centering
\includegraphics[scale=0.5]{./Figures/exampleastarperfectsolver}
%\decoRule
\caption[Examples]{a* perfect solver example}
\label{fig:exampleastarperfectsolver}
\end{figure}

\subsubsection{Deep Learning Heuristic}
\subsubsection{Deep Reinforcement Learning Heuristic}

### 2022-07-31 latex/Chapters/SlidingPuzzle.tex
update

@@

>>>>

\subsection{God numbers and hardest puzzles}

@@

\subsection{Manhattan heuristic}
In this section, I verify empirically that, as expected, the overhead of adding penalty in Manhattan++ for the linear constraint (which have all been precomputed and stored in a database) is more than compensated for by the reduction in nodes expansion. I have run my solver script for (n=2, m=5) in performance test mode, for both Manhattan and Manhattan++, with 250 randomly shuffled puzzles with \textit{nb\_shuffles} from 0 to 60 by increment of 5, as well as with \textit{nb\_shuffles = $\inf$}. The resulting run time and nodes expansions are as follows:

\begin{figure}[H]
\centering
\hspace*{-2.5cm}
\includegraphics[scale=0.45]{./Figures/25SPPerformanceManhattan}
%\decoRule
\caption[SP]{Manhattan vs Manhattan++}
\label{fig:25SPPerformanceManhattan}
\end{figure}

As can be seen, for low difficulty (up to \textit{nb\_shuffles = 20}), the node expansions are about the same in both cases, and the overhead of adding the linear constraints penalty increases the run time. However, for any non trivial case, Manhattan++ outperforms considerably (by a factor of about 2.5). For the 250 \textit{perfectly} shuffled instances, I got the following:

\begin{center}
\begin{tabular}{l*{7}{c}r}
                              & avg cost  & max cost & avg nodes & max nodes & avg run time (ms) & max run time (ms) \\
\hline
Manhattan                   &  34.7  & 50 & 133,332 & 2,110,887 & 49,561 & 606,838 \\
Manhattan++              & 34.7 &  50 & 53,637 & 962,324 & 19,723 & 239,468 \\
Improvement               & n/a &  n/a & x2.5 & x2.2 & x2.5 & x2.5 \\
\end{tabular}
\end{center}
